Report on the Third Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE3)

### 2015-11-05 WSSSPE3/report/WSSSPE3_review.tex
edits in lightning talk texts

\item \textbf{Birgit Penzenstadler, Colin Venters, Christoph Becker, Stefanie
Betz, Ruzanna Chitchyan, Letícia Duboc, Steve Easterbrook, Guillermo
Rodriguez-Navas and Norbert Seyff: \textit{Manifesting the Ghost of the Future:
Sustainability}}
The concept of sustainability has become a topic of interest in the field of computing, which is
evidenced by the increase in the number of events that focus on the topic - nevertheless, it isn’t
well understood yet. Penzenstadler argued that we often define sustainability too narrowly. Instead,
sustainability at its heart is a systemic concept and must be viewed from a range of different
dimensions including environmental, economic, individual, social and technical. She introduced the
Karlskrona Manifesto on Software Design, which distills knowledge from a broad range of related work
on the topic of sustainability into a set of (mis-)perceptions and principles. The manifesto does
not proclaim that there is an easy, one size fits all solution around the corner, but rather points
out that sustainability is a ``wicked problem'' and is often misunderstood.
Due to these misperceptions, even though sustainability's importance is increasingly recognized,
many software systems are unsustainable. Even more alarming is that most software systems'
broader impacts on sustainability are unknown. To change this, the Karlskrona Manifesto proposes
nine principles and commitments. These commitments are not dogmatic laws, but rather a commitment to
rethink, to move beyond the silo mentality, and analyze in more depth. As such, they do not
restrict, but rather open up a space for discussion.

@@

I argue that for scientific software to be truly sustainable, there is a need
for developers to use appropriate software engineering practices. My experience
interacting with scientific teams indicates that choosing and tailoring these
practices is not a trivial exercise. There is a general culture clash between
software engineering and science that hinders our ability to communicate and
choose appropriate methods. In addition, many experience scientific software
developers appear to be unaware of software engineering practices that may be
beneficial to them. The most appropriate software engineering practices are
those that are lightweight, properly tailored, and focus on the key software
development problems faced by scientists. In order to increase the use of
software engineering in science, we need more documented success stories. These
successes need to be socialized within the scientific community through
workshops like the Software Engineering for Science workshop
series~\footnote{\href{http://www.SE4Science.org/workshops}{http://www.SE4Science.org/workshops}}
and the new Software Engineering track in
Computing in Science and Engineering magazine.

>>>>

\item \textbf{Birgit Penzenstadler, Colin Venters, Christoph Becker, Stefanie
Betz, Ruzanna Chitchyan, Letícia Duboc, Steve Easterbrook, Guillermo
Rodriguez-Navas and Norbert Seyff: \textit{Manifesting the Ghost of the Future:
Sustainability}}
The concept of sustainability has become a topic of interest in the field of computing, which is
evidenced by the increase in the number of events that focus on the topic. Nevertheless, it isn't
well understood yet. Penzenstadler argued that we often define sustainability too narrowly. Instead,
sustainability at its heart is a systemic concept and must be viewed from a range of different
dimensions including environmental, economic, individual, social and technical. She introduced the
Karlskrona Manifesto on Software Design~\cite{Becker:2014}, which distills knowledge from a broad range of related work
on the topic of sustainability into a set of (mis-)perceptions and principles. The manifesto does
not proclaim that there is an easy, one-size-fits-all solution around the corner, but rather points
out that sustainability is a ``wicked problem'' and is often misunderstood.
Due to these misperceptions, even though sustainability's importance is increasingly recognized,
many software systems are unsustainable. Even more alarming is that most software systems'
broader impacts on sustainability are unknown. To change this, the Karlskrona Manifesto proposes
nine principles and commitments. These commitments are not dogmatic laws, but rather a commitment to
rethink, to move beyond the silo mentality, and analyze in more depth. As such, they do not
restrict, but rather open up a space for discussion.

@@

Jeff argued that for scientific software to be truly sustainable, there is a need
for developers to use appropriate software engineering practices. His experience
interacting with scientific teams indicates that choosing and tailoring these
practices is not a trivial exercise. There is a general culture clash between
software engineering and science that hinders our ability to communicate and
choose appropriate methods. In addition, many experience scientific software
developers appear to be unaware of software engineering practices that may be
beneficial to them. The most appropriate software engineering practices are
those that are lightweight, properly tailored, and focus on the key software
development problems faced by scientists. In order to increase the use of
software engineering in science, we need more documented success stories. These
successes need to be socialized within the scientific community through
workshops like the Software Engineering for Science workshop
series\footnote{\href{http://www.SE4Science.org/workshops}{http://www.SE4Science.org/workshops}}
and the new Software Engineering track in
\textit{Computing in Science and Engineering} magazine.
